### 2018-04-08 Writing/disparity_review.tex
updates on the structure

\section{Find a question} %NC: I wonder if "find a question" is exactly what we mean here. This suggests you've decided to use disparity whatever happens, and just need the right question. Maybe change to What is your question? 
Why are you performing a disparity analysis?
This seems like it should be obvious, but we stress its crucial importance.
The last decade has seen a vast array of new methods being published and implemented in freely available software, making disparity analyses easy to perform \citep{bouxin2005ginkgo,oksanen2007vegan,geiger2008,zelditch2012geometric,adams2013geomorph,Claddis,dispRityv02,adams2017geometric}.
However, this (consciously or not) pushes authors towards performing disparity analyses simply because of their ease of use, rather than stemming from a question driven approach.
Before rushing into a disparity analysis, one should consider some basic caveats.
For example, one major thing to consider is circularity: if the aim is to explore the variation in a dataset, disparity analysis is perfectly valid.
However, if this exploratory analysis reveals distinct groups, it would be circular to then test for differences among those groups as if this had been your hypothesis all along [CITE stats 1.0].
This can easily be avoided by using a subset of the data for preliminary data exploration, and then using the full dataset to test for differences among any groups that emerge.
Clearly framing your questions and their null expectations will help in designing better disparity analyses.

\section{Collect the appropriate data}
% If disparity is used to describe patterns that are explained by a certain process, every possible data will always only be a proxy for these patterns. % NC: Don't really understand your point here.
Once you have a clear question, it is important to collect the appropriate data for that question to avoid some major caveats.
For example, if a study focusses on the ecological adaptive radiation of a group, ecological traits are crucial and should be preferred over purely evolutionary traits.
Tangentially, more data are not necessarily better, especially when the data were originally collected for different questions (e.g. cladistic data - see below).

\subsection{Which types of data?} % NC: Not sure having these subheadings as Q looks best, maybe just Data types?
Morphological disparity analysis can be based on three main types of data:
\begin{enumerate}
    \item discrete data that can be either describing a feature (often referred to as ``cladistic''; e.g. the presence or absence of legs, the colour of a tail, etc. [CITE]) or counts (e.g. number of digits, etc. [CITE]);
    \item continuous data that can be either simple measurements (e.g. body length, etc. [CITE]) or more abstracted methods (e.g. geometric morphometrics, mathematical contour description, etc. [CITE]);
    \item or any mixture of discrete and continuous data (e.g. ratios [CITE]).
\end{enumerate}
None of these types of data are fundamentally better but each will be more appropriate for specific questions.
For example, when investigating variation within bat wing shapes, homologous landmarks (for geometric morphometrics analysis) or continuous measurements of bones may be appropriate.
On the other hand, if the question focuses on the convergence between bat and bird wings, the homology of the traits to measure is more complex to define and the contour of the wings might more readily answer the question. % NC: This section might need more explanation and/or some more examples

>>>>

\section{What is your question}
Why are you performing a disparity analysis?
This seems like it should be obvious, but we stress its crucial importance.
The last decade has seen a vast array of new methods being published and implemented in freely available software, making disparity analyses easy to perform \citep{bouxin2005ginkgo,oksanen2007vegan,geiger2008,zelditch2012geometric,adams2013geomorph,Claddis,dispRityv02,adams2017geometric}.
However, this (consciously or not) pushes authors towards performing disparity analyses simply because of their ease of use, rather than stemming from a question driven approach.
Before rushing into a disparity analysis, one should consider some basic caveats.
For example, one major thing to consider is circularity: if the aim is to explore the variation in a dataset, disparity analysis is perfectly valid.
However, if this exploratory analysis reveals distinct groups, it would be circular to then test for differences among those groups as if this had been your hypothesis all along [CITE stats 1.0].
This can easily be avoided by using a subset of the data for preliminary data exploration, and then using the full dataset to test for differences among any groups that emerge.
Clearly framing your questions and their null expectations will help in designing better disparity analyses.

\section{Collect the appropriate data}
Once you have a clear question, it is important to collect the appropriate data for that question to avoid some major caveats.
For example, if a study focusses on the ecological adaptive radiation of a group, ecological traits are crucial and should be preferred over purely evolutionary traits.
Tangentially, more data are not necessarily better, especially when the data were originally collected for different questions (e.g. cladistic data - see below).

\subsection{Data types}
Morphological disparity analysis can be based on three main types of data:
\begin{enumerate}
    \item discrete data that can be either describing a feature (often referred to as ``cladistic''; e.g. the presence or absence of legs, the colour of a tail, etc. [CITE]) or counts (e.g. number of digits, etc. [CITE]);
    \item continuous data that can be either simple measurements (e.g. body length, etc. [CITE]) or more abstracted methods (e.g. geometric morphometrics, mathematical contour description, etc. [CITE]);
    \item or any mixture of discrete and continuous data (e.g. ratios [CITE]).
\end{enumerate}
None of these types of data are fundamentally better but each will be more appropriate for specific questions.
For example, when investigating variation within bat wing shapes, homologous landmarks (for geometric morphometrics analysis) or continuous measurements of bones may be appropriate.
On the other hand, if the question focuses on the convergence between bat and bird wings, the homology of the traits to measure is more complex to define and the contour of the wings might more readily answer the question. % NC: This section might need more explanation and/or some more examples

%TG: add the bat example for above
